---
output: 
  html_document: 
    keep_md: yes
---
Reproducible Research: Course Project 1 Submission
==================================================



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1: Read and process dataset
Load packages, read and process data

```{r data}
library(dplyr)
library(lubridate)

# read data
amd <- read.csv("activity.csv", stringsAsFactors=FALSE)

# transform date column from character to date
amd$date <- mdy(amd$date)

# data grouped by date
amd %>% group_by(date) %>%
        summarize(steps = sum(steps)) -> amd2

# data grouped by interval
amd %>% na.omit() %>% 
        group_by(interval) %>%
        summarize(steps = mean(steps)) -> amd3

```


## 2) Histogram of the total number of steps taken each day


```{r daily steps, echo=TRUE}
library(ggplot2)

# histogram of the total number of steps each day
ggplot(amd2, aes(x=steps)) +
        geom_histogram(color="black", fill="blue")
```


## 3) Mean and median number of steps taken each day


```{r mean and median, echo=TRUE}
summary(amd2$steps)
```

## 4) Time series plot of the average number of steps taken

```{r time series, echo=TRUE}
# time series plot of average steps 
plot(amd3$interval, amd3$steps, type="l",xlab="5-Minute Interval",ylab="Average Number of Steps Taken")

```

## 5) The 5-minute interval that, on average, contains the maximum number of steps

```{r top interval, echo=TRUE}
amd3 %>% filter(print(steps), print(steps == max(steps)), steps == max(steps))
```

## 6) Code to describe and show a strategy for imputing missing data

Number of rows in the dataset with missing data:

```{r missing rows, echo=TRUE}
# calculate number of rows with missing values in the dataset
sum(!complete.cases(amd))
```

Code for imputing missing data with mean of 5-minute interval:

```{r impute means, echo=TRUE}
# impute mean (by 5-minute interval) for missing values in dataset
amd %>%
        group_by(interval) %>%
        mutate(steps = replace(steps, is.na(steps), mean(steps, na.rm = TRUE))) -> amd5

# new data grouped by date
amd5 %>% group_by(date) %>%
        summarize(steps = sum(steps)) -> amd6
```

## 7) Histogram of the total number of steps taken each day after missing values are imputed

```{r histogram of new data, echo=TRUE}
# histogram of the total number of steps each day with new dataset
ggplot(amd6, aes(x=steps)) +
        geom_histogram(color="black", fill="blue")
```

## 8) Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends


```{r panel plot, echo=TRUE}
# weekday dataset
amd5 %>% mutate(day = weekdays(date),
                daytype = if_else(day == "Monday"|day == "Tuesday"|day=="Wednesday"|day=="Thursday"|day=="Friday", "Weekday","Weekend")) -> daytype

# weekday dataset
daytype %>%
        filter(daytype == "Weekday") %>%
        group_by(interval) %>%
        summarize(steps = sum(steps)) -> weekday

#weekend dataset
daytype %>%
        filter(daytype == "Weekend") %>%
        group_by(interval) %>%
        summarize(steps = sum(steps)) -> weekend

# time series plot of average steps for weekdays
plot(weekday$interval, weekday$steps, type="l",xlab="5-Minute Interval",ylab="Average Number of Steps Taken")

# time series plot of average steps for weekdends
plot(weekend$interval, weekend$steps, type="l",xlab="5-Minute Interval",ylab="Average Number of Steps Taken")

```


